### 2019-04-11 analysis/analysis_report.rmd
added necessary functions and started report

---
title: "Analysis_Report"
author: "Daniel Pearson"
date: "11 April 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(afex)
library(emmeans)
library(here)
library(patchwork)
library(cowplot)
library(BayesFactor)
```