### 2019-04-18 README.Rmd
readme changes

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Overt attentional capture by reward-related stimuli overcomes top-down suppression"
```

# vmacfeaturesearch

This repository contains the data and code for our project investigating whether overt attention to reward-related distractors can be suppressed under feature search conditions in the same way as for physically salient distractors (see Gaspelin, Leonard, & Luck, 2017 - Attn Percept Psychophys).

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/danielpearson90/vmacfeaturesearch/archive/master.zip>

Or you can install this compendium as an R package, vmacfeaturesearch, from GitHub with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("danielpearson90/vmacfeaturesearch")
```
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/), Copyright (c) 2019 Daniel Pearson.

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/), Copyright (c) 2019 Daniel Pearson.