### 2019-05-14 README.Rmd
binder working

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Overt attentional capture by reward-related stimuli overcomes inhibitory suppression"
```

# vmacfeaturesearch

[![Travis-CI Build Status](https://travis-ci.org/danielpearson90/vmacfeaturesearch.svg?branch=master)](https://travis-ci.org/danielpearson90/vmacfeaturesearch) [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/danielpearson90/vmacfeaturesearch/master/?urlpath=rstudio)

This repository contains the data and analysis code for our project investigating whether overt attention to reward-related distractors can be suppressed under feature search conditions.

Our pre-print is online here:

> Pearson, D., Watson, P., Cheng, P., & Le Pelley, M. (2019). _`r Title`_. PsyArXiv, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.31234/osf.io/db9hg>

### How to cite

Please cite this compendium as:

> Pearson, D., Watson, P., Cheng, P., & Le Pelley, M. (`r format(Sys.Date(), "%Y")`). Investigating attentional suppression of reward-related distractors. Retrieved from <https://osf.io/yrdzv>

### Contents
- [:file_folder: analysis](/analysis): Contains analysis report in R Markdown format, as well as R script format.
- [:file_folder: analysis/paper](/analysis/paper): R markdown source document for the manuscript.
- [:file_folder: analysis/data](/analysis/data): Data files that have been dericed from the raw eye-tracker data. The raw eye-gaze files are very large and so are not stored here. They can be found at: <https://osf.io/yrdzv/>.
- [:file_folder: analysis/figures](/analysis/figures): Figures generated in the analysis.

### Reproducibility
This repository is organized as a reproducible research compendium. 
Click the ![Binder](https://mybinder.org/badge_logo.svg) button above to explore in an interactive RStudio session (this takes approximate 15-20 mins to build). Binder uses [rocker-project.org](https://rocker-project.org) Docker images to ensure a consistent and reproducible computational environment.  These Docker images can also be used locally.

To explore the code locally, clone or download this repository into RStudio or your perferred environment and install the compendium by running `devtools::install(dep=TRUE)`.

This compendium is checked by Travis-CI continuous integration.  Click the ![Build Status](https://travis-ci.org/cboettig/noise-phenomena.svg?branch=master) button above for details.

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/), Copyright (c) 2019 Daniel Pearson.

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/), Copyright (c) 2019 Daniel Pearson.